---
title: "Reproducibility Report: Group B Choice 3"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

-------

For this exercise, please try to reproduce the results from Experiment 2 of the associated paper (de la Fuente, Santiago, Roman, Dumitrache, & Casasanto, 2014). The PDF of the paper is included in the same folder as this Rmd file. 

#### Methods summary: 

Researchers tested the question of whether temporal focus differs between Moroccan and Spanish cultures, hypothesizing that Moroccans are more past-focused, whereas Spaniards are more future-focused. Two groups of participants ($N = 40$ Moroccan and $N=40$ Spanish) completed a temporal-focus questionnaire that contained questions about past-focused ("PAST") and future-focused ("FUTURE") topics. In response to each question, participants provided a rating on a 5-point Likert scale on which lower scores indicated less agreement and higher scores indicated greater agreement. The authors then performed a mixed-design ANOVA with agreement score as the dependent variable, group (Moroccan or Spanish, between-subjects) as the fixed-effects factor, and temporal focus (past or future, within-subjects) as the random effects factor. In addition, the authors performed unpaired two-sample t-tests to determine whether there was a significant difference between the two groups in agreement scores for PAST questions, and whether there was a significant difference in scores for FUTURE questions.

------

#### Target outcomes: 

Below is the specific result you will attempt to reproduce (quoted directly from the results section of Experiment 2):

> According to a mixed analysis of variance (ANOVA) with group (Spanish vs. Moroccan) as a between-subjects factor and temporal focus (past vs. future) as a within-subjectS factor, temporal focus differed significantly between Spaniards and Moroccans, as indicated by a significant interaction of temporal focus and group, F(1, 78) = 19.12, p = .001, ηp2 = .20 (Fig. 2). Moroccans showed greater agreement with past-focused statements than Spaniards did, t(78) = 4.04, p = .001, and Spaniards showed greater agreement with future-focused statements than Moroccans did, t(78) = −3.32, p = .001. (de la Fuente et al., 2014, p. 1685).

------

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

# Step 1: Load packages

```{r}
library(tidyverse) # for data munging
library(knitr) # for kable table formating
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files
library(afex) # anova functions
library(ez) # anova functions 2
library(scales) # for plotting
library(Rmisc)
library(rstatix)
std.err <- function(x) sd(x)/sqrt(length(x)) # standard error
```

# Step 2: Load data

```{r}
# Just Experiment 2
setwd("~/Desktop/psych251_problem_sets/ps3/Group B/Choice 3/data")
data_path <- 'DeLaFuenteEtAl_2014_RawData.xls'
d <- read_excel(data_path, sheet=3)
```

# Step 3: Tidy data

```{r}

names(d)
colnames(d) <- c('group', 'participant', 'subscale', 'item', 'agreement')
d$agreement <- sapply(d$agreement, as.numeric)
#View(d)

describe_d <- d %>%
  group_by(participant, group) %>%
  dplyr::summarize(n())
#View(describe_d)

d <- filter(d, participant != 40, participant != 25, participant != 24) 

d_wide <- pivot_wider(data = d, names_from = item, values_from = agreement)
colnames(d_wide) <- c('group', 'participant', 'subscale', 'Q1', 'Q2', 'Q3', 'Q4', 'Q5', 'Q6', 'Q7', 'Q8', 'Q9', 'Q10', 'Q11', 'Q12', 'Q13', 'Q14', 'Q15', 'Q16', 'Q17', 'Q18', 'Q19', 'Q20', 'Q21')
d_wide <- d_wide %>%
  mutate(average_agreement = rowMeans(d_wide[ , c('Q1', 'Q2', 'Q3', 'Q4', 'Q5', 'Q6', 'Q7', 'Q8', 'Q9', 'Q10', 'Q11', 'Q12', 'Q13', 'Q14', 'Q15', 'Q16', 'Q17', 'Q18', 'Q19', 'Q20', 'Q21')], na.rm = T))
#View(d_wide)

```

# Step 4: Run analysis

## Pre-processing

```{r}

# I am not sure what to do here? 

```

## Descriptive statistics

Try to recreate Figure 2 (fig2.png, also included in the same folder as this Rmd file):

```{r}

names(d_wide)

d_summary <- summarySE(d_wide, measurevar = 'average_agreement', groupvars = c('group', 'subscale'), na.rm = TRUE)
d_summary

ggplot(data = d_summary, 
       aes(y = average_agreement, x = group, fill = subscale)) +
       geom_bar(position = position_dodge(), stat = 'identity') +
       geom_errorbar(aes(ymin = average_agreement - se, ymax = average_agreement + se), width = .3, position = position_dodge(.9)) +
       labs(x = "Group", y = "Rating")

```

## Inferential statistics

> According to a mixed analysis of variance (ANOVA) with group (Spanish vs. Moroccan) as a between-subjects factor and temporal focus (past vs. future) as a within-subjects factor, temporal focus differed significantly between Spaniards and Moroccans, as indicated by a significant interaction of temporal focus and group, F(1, 78) = 19.12, p = .001, ηp2 = .20 (Fig. 2).

```{r}
# reproduce the above results here

anovaresult <- aov(average_agreement ~ group*subscale, data = d_wide)
summary(anovaresult)

```

> Moroccans showed greater agreement with past-focused statements than Spaniards did, t(78) = 4.04, p = .001, 

```{r}
# reproduce the above results here

d_wide_past <- d_wide %>%
  filter(subscale == 'PAST')

t.test(average_agreement ~ group, data = d_wide_past)

```

> and Spaniards showed greater agreement with future-focused statements than Moroccans did, t(78) = −3.32, p = .001.(de la Fuente et al., 2014, p. 1685)

```{r}
# reproduce the above results here

d_wide_future <- d_wide %>%
  filter(subscale == 'FUTURE')

t.test(average_agreement ~ group, data = d_wide_future)

```

# Step 5: Reflection

Were you able to reproduce the results you attempted to reproduce? If not, what part(s) were you unable to reproduce?

> No, I think I did something wrong with data wrangling because I couldn't figure out how to proceed unless I got rid of the three atypical participants from the dataset, but I wanted to finish the rest of the assignment so I just proceeded with deleting the three participants, and I think that made my result look different from the original paper 

How difficult was it to reproduce your results? 

> Difficult, I was having a hard time with dealing with the three atypical responses

What aspects made it difficult? What aspects made it easy?

> I still don't know what was the best way to deal with the three atypical responses
